Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Clustering and Parameter Estimation

### 2020-10-21 summary.Rmd
more simulations

\begin{algorithm}[H]
  \DontPrintSemicolon
  \SetAlgoLined
  \caption{PABM parameter estimation using the adjacency matrix.}
  \KwData{Adjacency matrix $A$, community assignments $1, ..., K$}
  \KwResult{PABM parameter estimates $\{\hat{\lambda}^{(kl)}\}_K$.}
  Arrange the rows and columns of $A$ by community such that each 
  $A^{(kl)}$ block consists of estimated edge probabilities between 
  communities $k$ and $l$.\;
  \For{$k, l = 1, ..., K$, $k \leq l$} {
    Compute $A^{(kl)} = U \Sigma V^\top$, the SVD of the $kl^{th}$ 
    block.\;
    Assign $u^{(kl)}$ and $v^{(kl)}$ as the first columns of $U$ and $V$. 
    Assign $\sigma^2 \leftarrow \Sigma_{11}$.\;
    Assign $\hat{\lambda}^{(kl)} \leftarrow \pm \sigma^{(kl)} u^{(kl)}$ and 
    $\hat{\lambda}^{(lk)} \leftarrow \pm \sigma^{(kl)} v^{(kl)}$.
  }
\end{algorithm}

@@

```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against modularity maximization (blue) and sparse subspace clustering (green) for sample sizes from 128 to 4096 and number of underlying communities from 2 to 4. Simulations were repeated 40 times for each sample size. The $y$-axis is shown on a logarithmic scale.', fig.width = 10}
clustering.df <- readr::read_csv('clustering-k.csv')
clustering.df %>%
  dplyr::group_by(K, n) %>%
  dplyr::summarise(med.err = median(error),
                   first.q = quantile(error, .25),
                   third.q = quantile(error, .75),
                   med.err.ssc = median(error.ssc),
                   first.q.ssc = quantile(error.ssc, .25),
                   third.q.ssc = quantile(error.ssc, .75),
                   med.err.mm = median(error.mm),
                   first.q.mm = quantile(error.mm, .25),
                   third.q.mm = quantile(error.mm, .75)) %>%
  dplyr::ungroup() %>% 
  ggplot() +
  # scale_x_log10() + 
  labs(y = 'community detection error', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err,
                colour = 'Algorithm 2')) +
  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
                    colour = 'Algorithm 2')) + 
  geom_line(aes(x = n, y = med.err.ssc,
                colour = 'SSC')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
                    colour = 'SSC')) + 
  geom_line(aes(x = n, y = med.err.mm,
                colour = 'modularity maximization')) +
  geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
                    colour = 'modularity maximization')) +
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
  scale_y_log10()
```

@@

```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 128 to 4096. Simulations were repeated 40 times for each sample size. The $y$-axis is shown on a logarithmic scale.', cache = FALSE, fig.width = 10}
clustering.df <- readr::read_csv('clustering-k.csv')
clustering.df %>%
  dplyr::group_by(n, K) %>%
  dplyr::summarise(med.err = median(error),
                   first.q = quantile(error, .25),
                   third.q = quantile(error, .75),
                   med.err.ssc = median(error.ssc),
                   first.q.ssc = quantile(error.ssc, .25),
                   third.q.ssc = quantile(error.ssc, .75)) %>%
  dplyr::ungroup() %>% 
  ggplot() +
  # scale_x_log10() + 
  labs(y = 'misclustering count', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err * n,
                colour = 'Algorithm 2')) +
  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
                    colour = 'Algorithm 2')) + 
  geom_line(aes(x = n, y = med.err.ssc * n,
                colour = 'SSC')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
                    colour = 'SSC')) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_y_log10() +
  scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096))
```

```{r lambda_est_k, cache = FALSE, fig.cap = 'Log-log plot of the median and IQR RMSE from Algorithm 4 (red) compared against an MLE-based method (blue) using sample sizes from 256 to 4096 and number of clusters from 2 to 4. Simulations were repeated 50 times for each sample size.', fig.width = 10}
rmse.df <- readr::read_csv('rmse.csv') %>% 
  dplyr::transmute(K = 2,
                   n = n,
                   rmse = rmse,
                   rmse.mle = rmse.mle) %>% 
  dplyr::bind_rows(readr::read_csv('rmse-k.csv'))

rmse.df %>% 
  na.omit() %>% 
  dplyr::group_by(K, n) %>% 
  dplyr::summarise(median.rmse = median(rmse),
                   q1.rmse = quantile(rmse, .25),
                   q3.rmse = quantile(rmse, .75),
                   median.rmse.mle = median(rmse.mle),
                   q1.rmse.mle = quantile(rmse.mle, .25),
                   q3.rmse.mle = quantile(rmse.mle, .75)) %>% 
  dplyr::ungroup() %>% 
  ggplot() +
  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 4')) + 
  geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
                    colour = 'Algorithm 4')) + 
  geom_line(aes(x = n, y = median.rmse.mle, colour = 'MLE-based')) + 
  geom_errorbar(aes(x = n, ymin = q1.rmse.mle, ymax = q3.rmse.mle,
                   colour = 'MLE-based')) + 
  scale_colour_brewer(palette = 'Set1') + 
  labs(y = 'RMSE', colour = NULL) + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_y_log10() + 
  scale_x_log10(breaks = c(64, 128, 256, 512, 1024, 2048))
```

>>>>

\begin{algorithm}[H]
  \DontPrintSemicolon
  \SetAlgoLined
  \caption{PABM parameter estimation using the adjacency matrix.}
  \KwData{Adjacency matrix $A$, community assignments $1, ..., K$}
  \KwResult{PABM parameter estimates $\{\hat{\lambda}^{(kl)}\}_K$.}
  Arrange the rows and columns of $A$ by community such that each 
  $A^{(kl)}$ block consists of estimated edge probabilities between 
  communities $k$ and $l$.\;
  \For{$k, l = 1, ..., K$, $k \leq l$} {
    Compute $A^{(kl)} = U \Sigma V^\top$, the SVD of the $kl^{th}$ 
    block.\;
    Assign $u^{(kl)}$ and $v^{(kl)}$ as the first columns of $U$ and $V$. 
    Assign $(\sigma^{(kl)})^2 \leftarrow \Sigma_{11}$.\;
    Assign $\hat{\lambda}^{(kl)} \leftarrow \pm \sigma^{(kl)} u^{(kl)}$ and 
    $\hat{\lambda}^{(lk)} \leftarrow \pm \sigma^{(kl)} v^{(kl)}$.
  }
\end{algorithm}

@@

```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against modularity maximization (blue) and sparse subspace clustering (green) for sample sizes from 128 to 4096 and number of underlying communities from 2 to 4. Simulations were repeated 40 times for each sample size. The $y$-axis is shown on a logarithmic scale.', fig.width = 10}
clustering.df <- readr::read_csv('clustering-k.csv')
clustering.df %>%
  dplyr::group_by(K, n) %>%
  dplyr::summarise(med.err = median(error),
                   first.q = quantile(error, .25),
                   third.q = quantile(error, .75),
                   med.err.ssc = median(error.ssc),
                   first.q.ssc = quantile(error.ssc, .25),
                   third.q.ssc = quantile(error.ssc, .75),
                   med.err.mm = median(error.mm),
                   first.q.mm = quantile(error.mm, .25),
                   third.q.mm = quantile(error.mm, .75)) %>%
  dplyr::ungroup() %>% 
  ggplot() +
  # scale_x_log10() + 
  labs(y = 'community detection error', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err,
                colour = 'Algorithm 2')) +
  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
                    colour = 'Algorithm 2')) + 
  geom_line(aes(x = n, y = med.err.ssc,
                colour = 'SSC')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
                    colour = 'SSC')) + 
  geom_line(aes(x = n, y = med.err.mm,
                colour = 'Modularity maximization')) +
  geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
                    colour = 'Modularity maximization')) +
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
  scale_y_log10()
```

@@

```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 128 to 4096. Simulations were repeated 40 times for each sample size. The $y$-axis is shown on a logarithmic scale.', cache = FALSE, fig.width = 10}
clustering.df <- readr::read_csv('clustering-k.csv')
clustering.df %>%
  dplyr::group_by(n, K) %>%
  dplyr::summarise(med.err = median(error),
                   first.q = quantile(error, .25),
                   third.q = quantile(error, .75),
                   med.err.ssc = median(error.ssc),
                   first.q.ssc = quantile(error.ssc, .25),
                   third.q.ssc = quantile(error.ssc, .75),
                   med.err.mm = median(error.mm),
                   first.q.mm = quantile(error.mm, .25),
                   third.q.mm = quantile(error.mm, .75)) %>%
  dplyr::ungroup() %>% 
  ggplot() +
  # scale_x_log10() + 
  labs(y = 'misclustering count', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err * n,
                colour = 'Algorithm 2')) +
  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
                    colour = 'Algorithm 2')) + 
  geom_line(aes(x = n, y = med.err.ssc * n,
                colour = 'SSC')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
                    colour = 'SSC')) + 
  geom_line(aes(x = n, y = med.err.mm * n,
                colour = 'Modularity maximization')) + 
  geom_errorbar(aes(x = n, ymin = first.q.mm * n, ymax = third.q.mm * n,
                    colour = 'Modularity maximization')) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_y_log10() +
  scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096))
```

```{r lambda_est_k, cache = FALSE, fig.cap = 'Log-log plot of the median and IQR RMSE from Algorithm 4 (red) compared against an MLE-based method (blue) using sample sizes from 128 to 4096 and number of clusters from 2 to 4. Simulations were repeated 50 times for each sample size.', fig.width = 10}
rmse.df <- readr::read_csv('rmse-k.csv')

rmse.df %>% 
  na.omit() %>% 
  dplyr::group_by(K, n) %>% 
  dplyr::summarise(median.rmse = median(rmse),
                   q1.rmse = quantile(rmse, .25),
                   q3.rmse = quantile(rmse, .75),
                   median.rmse.mle = median(rmse.mle),
                   q1.rmse.mle = quantile(rmse.mle, .25),
                   q3.rmse.mle = quantile(rmse.mle, .75)) %>% 
  dplyr::ungroup() %>% 
  ggplot() +
  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 4')) + 
  geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
                    colour = 'Algorithm 4')) + 
  geom_line(aes(x = n, y = median.rmse.mle, colour = 'MLE-based')) + 
  geom_errorbar(aes(x = n, ymin = q1.rmse.mle, ymax = q3.rmse.mle,
                   colour = 'MLE-based')) + 
  scale_colour_brewer(palette = 'Set1') + 
  labs(y = 'RMSE', colour = NULL) + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_y_log10() + 
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096))
```
